---
title: "Survey Methodology II - Final Project"
author: "Eric Hausken, Alicia Talavera, Giorgio Rivoli"
format: html
editor: visual
---

From 2012 to 2019, surveys were conducted at the European Union level focusing on aspects such as discrimination existing in the EU. In 2019, one of the questions was: *"Do you think that transgender people should be able to change their civil documents to match their inner gender identity?"* The objective of this work is to detect which variables influence whether an individual responds positively or negatively to this question.

## Packages used

```{r}
library(tidyverse)
library(haven)
library(readxl)
library(mice)
library(ggplot2)
library(gridExtra)
library(caret)
```

# Loading databases

Our main database is the EU survey conducted in 2019. Then, we added some variables on country level. First, a GINI index database, which shows the value of the index for each country. Second, more specifically about LGBT+ rights in the EU countries, a database which includes data regarding whether a country has legal homosexual marriage, procedures for trans people name or legal sex change, hate crime laws, etc. This database is from ILGA-Europe and the Rainbow Europe Map and Index 2019.

```{r}
# Main database
# data <- read_dta("ZA7575.dta")
data <- read_dta("Data/ZA7575.dta")

# GINI index database
gini <- read_csv("eu_gini_coef.csv")

# LGBTphobia database
ilga <- read_xlsx("ilga.xlsx", sheet = 2)
ilga <- ilga |> 
  filter(CONTINENT == "EUROPE")

```

### Merging together Countries database

Now, we merge both GINI and ILGA databases together to create a 'countries' database.

```{r}

names <- data.frame(
  COUNTRY = c(
    "Belgium", "Greece", "Lithuania", "Portugal", "Bulgaria", "Spain", "Luxembourg", "Romania", "Czechia", "France", 
    "Hungary", "Slovenia", "Denmark", "Croatia", "Malta", "Slovakia", "Germany", "Italy", "Netherlands", "Finland", "Estonia", "Cyprus", "Austria", "Sweden", "Ireland", "Latvia", "Poland",  "Iceland", "Norway", "Liechtenstein", "Switzerland",
  "Bosnia and Herzegovina", "Montenegro", "Moldova", "North Macedonia", "Georgia", "Albania", "Serbia", "Türkiye", 
    "Ukraine","Kosovo"
  ),
  geo = c(
    "BE", "EL", "LT", "PT", "BG", "ES", "LU", "RO", "CZ", "FR", "HU", "SI", "DK", "HR", "MT", "SK", "DE", "IT", "NL", "FI", "EE", "CY", "AT", "SE", "IE", "LV", "PL", "IS", "NO", "LI", "CH", "BA", "ME", "MD", "MK", "GE", "AL", "RS", "TR", "UA","XK"
  )
)

gini <- gini |> 
  select(geo, TIME_PERIOD, OBS_VALUE, OBS_FLAG) |> 
    filter(TIME_PERIOD == "2019") |> 
  left_join(names, by = "geo")


countries <- gini |> 
  left_join(ilga, by = "COUNTRY") |> 
  select(-TIME_PERIOD, CONTINENT) |> 
  filter(!is.na(CONTINENT))

# Turn all ILGA variables to factors
countries <- countries |> 
    mutate_at(vars(7:ncol(countries)), as.factor)
  
```

## Selecting variables in the main dataset

There are over 600 columns in the original database, so we selected only a few that were relevant to the topic, as well as sociodemographic variables.

```{r}

selected_data <- data |> 
  select(
    uniqid, 
    isocntry, # nationality keep 
    d11, # age  
    d70, # life satisfaction -recode below DK == NA
    polintr, # political interest -recode below DK == NA
    qa1, # globalization scale 1-4, convert 5 to NA 
    # qa5a,
    # qa7,
    # qa8,
    # qa9,
    # qa11,
    # qa12,
    # qa13,
    # qa14,
    # qa17,
    sd1_7, # do you know someone who is transgender. 
    # sd2_1, sd2_2, sd2_3, sd2_4, sd2_5, sd2_6, sd2_7, sd2_8,
    sd3, 
    qc1_4, 
    qc1_8,
    qc2_1:qc2_14, # recode to 1 for discrimination == yes
      qc2_15, # recode to 0. discriminatino == no
      qc2_16, # recode to NA, impute missing data.  DONT KNOW
    # qc3,
    # qc4,
    qc5_1, qc5_3,
    qc6_1r, qc6_2r, qc6_3r, qc6_10r, # is there discrimination towards...
    qc7,
    qc9_1,qc9_4, qc9_5, qc9_10,
    qc12_1r, qc12_2r, qc12_6r, qc12_10r, qc12_11r,
   # qc13_1r, qc13_2r, qc13_3r, qc13_4r, qc13_5r, qc13_6r, qc13_7r, qc13_8r, qc13_9r, qc13_10r, qc13_11r, qc13_12r, qc13_13r, qc13_14r, qc13_15r,
    qc15_1:qc15_3,
    qc17_1:qc17_5,
    qc18_1r, qc18_2r, qc18_3r,
    qc19, # TARGET VARIABLE 
    qc20,
    d1, # keep, scale from 0-10, remove 11, 12
    # d7,
    d10, # gender keep 
    d8, # keep education years 
    d15a, # convert to factor. occupation 
    d25, # keep type of city 
    # d40a, d40b, d40c, d43a, d43b,
    d60, # financial keep 
    d62_1,  # keep internet at home. calculate average 
    d63, # class keep. as factor 
    d72_1, d72_2, # keep. does my voice count
    p3r, # paradata
    p5, # paradata
    p4 # paradata
)



selected_data <- selected_data |> 
  rename(geo = isocntry) |> 
  left_join(names, by = "geo")
```

## Cleaning missing data

Regarding missing data, we drop out some variables for having too many NAs and impute the rest of varibles we will use. We decide not to impute the target variable to check how NAs distribution change.

```{r, fig.height=6, fig.width=12}

imputed_data <- selected_data |> 
    mutate(d70 = ifelse(d70 == 5, NA, as.numeric(d70))) |> 
    mutate(qa1 = ifelse(qa1 > 5, NA, as.numeric(qa1))) |> 
    mutate(polintr = ifelse(polintr == 4, NA, as.numeric(polintr))) |> 
    mutate(d63 = ifelse(d63 > 5, NA, as.numeric(d63))) |> 
    mutate(d72_1 = ifelse(d72_1 %in% c(5, 6), NA, as.numeric(d72_1))) |> 
    mutate(d72_2 = ifelse(d72_2 %in% c(5,6), NA, as.numeric(d72_2))) |> 
    mutate(sd3 = ifelse(sd3 %in% c(14,15,16), NA, as.numeric(sd3))) |> 
    mutate(qa1 = ifelse(qa1 %in% 5, NA, as.numeric(qa1))) |> 
    mutate(sd1_7 = ifelse(sd1_7 %in% c(3,4), NA, as.numeric(sd1_7))) |> 
    mutate(d25 = ifelse(d25 > 3, NA, as.numeric(d25))) |> 
    mutate(d8 = ifelse(d8 %in% c(98,99, 97), NA, as.numeric(d8))) |> 
   mutate(qc1_4 = ifelse(qc1_4 > 4, NA, as.numeric(qc1_4))) |> 
   mutate(qc1_8 = ifelse(qc1_8 > 4, NA, as.numeric(qc1_8))) |> 
   mutate(qc6_1r = ifelse(qc6_1r> 3, NA, as.numeric(qc6_1r))) |> 
   mutate(qc6_2r = ifelse(qc6_2r > 3, NA, as.numeric(qc6_2r))) |>    mutate(qc6_3r = ifelse(qc6_3r>3, NA, as.numeric(qc6_3r))) |> 
   mutate(qc6_10r = ifelse(qc6_10r>3,NA,as.numeric(qc6_10r))) |> 
   mutate(qc7 = ifelse(qc7>10,NA,as.numeric(qc7))) |> 
   mutate(qc9_1 = ifelse(qc9_1>4,NA,as.numeric(qc9_1))) |> 
   mutate(qc9_4 = ifelse(qc9_4>4,NA,as.numeric(qc9_4))) |> 
   mutate(qc9_5 = ifelse(qc9_5>4,NA,as.numeric(qc9_5))) |> 
   mutate(qc9_10 = ifelse(qc9_10>4,NA,as.numeric(qc9_10))) |> 
   mutate(qc5_1 = ifelse(qc5_1>2,NA,as.numeric(qc5_1))) |> 
   mutate(qc5_3 = ifelse(qc5_3>2,NA,as.numeric(qc5_3))) |> 
   mutate(qc12_1r = ifelse(qc12_1r>10,NA,as.numeric(qc12_1r))) |>
   mutate(qc12_2r = ifelse(qc12_2r>10,NA,as.numeric(qc12_2r))) |>
   mutate(qc12_6r = ifelse(qc12_6r>10,NA,as.numeric(qc12_6r))) |>
   mutate(qc12_10r = ifelse(qc12_10r>10,NA,as.numeric(qc12_10r))) |>
   mutate(qc12_11r = ifelse(qc12_11r>10,NA,as.numeric(qc12_11r))) |> 
   mutate(qc15_1 = ifelse(qc15_1>4,NA,as.numeric(qc15_1))) |> 
   mutate(qc15_2 = ifelse(qc15_2>4,NA,as.numeric(qc15_2))) |> 
   mutate(qc15_3 = ifelse(qc15_3>4,NA,as.numeric(qc15_3))) |> 
   mutate(qc17_1 = ifelse(qc17_1>4,NA,as.numeric(qc17_1))) |> 
   mutate(qc17_2 = ifelse(qc17_2>4,NA,as.numeric(qc17_2))) |> 
   mutate(qc17_3 = ifelse(qc17_3>4,NA,as.numeric(qc17_3))) |> 
   mutate(qc17_4 = ifelse(qc17_4>4,NA,as.numeric(qc17_4))) |> 
   mutate(qc17_5 = ifelse(qc17_5>4,NA,as.numeric(qc17_5))) |> 
   mutate(qc18_1r = ifelse(qc18_1r>3,NA,as.numeric(qc18_1r))) |> 
   mutate(qc18_2r = ifelse(qc18_2r>3,NA,as.numeric(qc18_2r))) |> 
   mutate(qc18_3r = ifelse(qc18_3r>3,NA,as.numeric(qc18_3r))) |> 
   mutate(qc20 = ifelse(qc20>2,NA,as.numeric(qc20))) |> 
   mutate(d1 = ifelse(d1>10,NA,as.numeric(d1)))

imputed_data <- imputed_data |> select("d70", "qa1", "polintr", "d63", "d72_1", "d72_2", "sd3", "sd1_7", "d25", "d8", "qc1_4", "qc1_8", "qc6_1r", "qc6_2r","qc6_3r", "qc6_10r", "qc7", "qc9_1", "qc9_4", "qc9_5","qc9_10", "qc5_1", "qc5_3", "qc12_1r", "qc12_2r", "qc12_6r", "qc12_10r", "qc12_11r", "qc15_1", "qc15_2", "qc15_3", "qc17_1", "qc17_2", "qc17_3", "qc17_3", "qc17_5", "qc18_1r", "qc18_2r", "qc18_3r", "qc20", "d1")

set.seed(123)

meth <- rep("rf", 40)

meth[c("d70", "qa1", "polintr", "d63", "d72_1", "d72_2", "sd3", "sd1_7", "d25", "d8", "qc1_4", "qc1_8", "qc6_1r", "qc6_2r","qc6_3r", "qc6_10r", "qc7", "qc9_1", "qc9_4", "qc9_5","qc9_10", "qc5_1", "qc5_3", "qc12_1r", "qc12_2r", "qc12_6r", "qc12_10r", "qc12_11r", "qc15_1", "qc15_2", "qc15_3", "qc17_1", "qc17_2", "qc17_3", "qc17_3", "qc17_5", "qc18_1r", "qc18_2r", "qc18_3r", "qc20", "d1")] <- "rf"

imputed_rf = mice(imputed_data, methid=meth, iterations = 10)

  
imputed_data <- complete(imputed_rf)


# Replace non imputed columns with imputed ones
imputed_vars <- names(imputed_data)

selected_data <- selected_data[, !(names(selected_data) %in% imputed_vars)]

selected_data <- cbind(selected_data, imputed_data)

colSums(is.na(selected_data))
```

## Recode variables

We recode many variables as factors and label them to be able to analyse them.

```{r}
recoded_data <- selected_data |> 
  mutate(
    
    qc19 = factor(
      case_when(
        qc19 == 2 ~ 0, # make No as zero 
        qc19 == 3 ~ NA, # make DK as NA
        qc19 == 1 ~ 1 # keep Yes as 1        
        ),
      # label the values 
      labels = c("No", "Yes")
      ), 
    
    qa1 = factor(qa1, 
                 labels = c("A lot", "Somewhat", "Not really", "Not at all")),
    qc2_sum = qc2_1 + qc2_2 + qc2_3 + qc2_4 +
              qc2_5 + qc2_6 + qc2_7 + qc2_8 +
              qc2_9 + qc2_10 + qc2_11 + qc2_12 +
              qc2_13 + qc2_14,
    qc2_recode = case_when(
      qc2_sum > 0 ~ 1,
      qc2_15 == 1 ~ 0,
      qc2_16 == 1 ~ NA),
    # qc6_recode  = mean(qc6_1, qc6_2),
    d15a = factor(d15a, ordered = FALSE ),
    d1 = replace(d1, d1 %in% c(11,12), NA),
    d70 = factor(d70, ordered = F,
          labels = c("Very", "Somewhat", "Not very", "Not at all")), 
    polintr = factor(polintr, ordered = F,
          labels = c("frequently", "occasionally", "never")),
    d63 = factor(d63, ordered = FALSE, labels = c("Working class", "Lower middle class", "Middle class", "Upper middle class", "Higher class")),
    d72_1 = factor(d72_1, ordered = FALSE, labels= c("Totally agree", "Agree", "Disagree", "Totally disagree")),
    d72_2 = factor(d72_1, ordered = FALSE, labels= c("Totally agree", "Agree", "Disagree", "Totally disagree")),
    sd3 = factor(sd3, ordered = FALSE, labels= c("Catholic", "Orthodox Christian", "Prostetant", "Other Christian", "Jewish", "Muslim - Shia", "Muslin - Sunni", "Other Muslim", "Sikh", "Buddhist", "Hindu", "Atheist", "Agnostic")),
    sd1_7 = factor(sd1_7, ordered = F, labels = c("Yes", "No")),
    d25 = factor(d25, ordered = F, labels = c("Rural area", "Small/medium town", "Large town")),
    age_group = cut(d11, breaks = seq(0, 100, by = 10), include.lowest = TRUE),
    school_group = cut(d8, breaks = seq(1, 50, by = 5), include.lowest = TRUE),
    qc20 = factor(qc20, ordered = F, labels = c("Yes", "No"))) 

# Join recoded data with countries data
recoded_data <- recoded_data |> 
  left_join(countries, by = "geo")

```

## Descriptive analysis

### Individual characteristics

Most people answer "yes" to our target question (52.7%), but there are also a significant amount of 'no' (35%) and 'dont know' answers (11.9%).

```{r}
prop.table(table(recoded_data$qc19, useNA = "ifany")) * 100

```

Now, we will visualize some plots crossing the answers in QC19 to other variables.

-   age_group- Age. In younger ages more people tend to answer 'yes', which means, more people are in favor of allowing trans people to change their documents. As the age increases, the 'yes' answer is less likely, but still over 50%. The number of NAs also increase with age, with a peak between 60 and 70 years old.

-   sd3 - Religion. La mayoría de personas de la muestra se consideran católicas. De éstas, un 56% está a favor frente un 43% en contra. La identidad en la cual hay más personas en contra, con un 57%, es orthodox christian. In the other hand, there are more individuals who agree who are Agnostic and atheists (Sikh too but there are very few cases in the sample, it's not representative).

-   sd1_7 - Friend/acquaintance who is trans. A 89.8% of europeans don't have a trans friend or aqcuaintance. However, from the people who don't know a trans person, 80% of them are in favor of trans people changing their document, a much higher percentage than the 57% between the people who do know a trans person.

-   qc2_recode - Suffered discrimination. Most people haven't suffered discrimination. The amount of people who is for and against the target question is very similar in both groups.

-   polintr - Political interest. More people who never get involved in political topics answered 'don't know' proportionally than in the rest of categories. Thy're also the individuals with a higher rate of 'no' answers. There are more people who answer 'no' and have a frequent interest in politics compared to people who have an ocasional interest.

-   qa1 - Opinion on international trade. People who consider they benefit a lot from international trade allowed by the EU tend to be more positive regarding our target variable. However, as the perception of trade decreases, so does the amount of people who are in favor, as well as people who don't answer.

-   d25 - Type of city. There is more people against trans people changing their documents in rural areas, as well as in large towns compared to medium-sized cities.

-   d10 - Gender. More women than men answered 'yes' in the target variable.

-   school_group- Age when stopped education. This variable can be colineal with age. People who abandoned education very young are less suportive. The older the age, the more the support.

-   d63 - Subjective social class. A 27% of people define themselves as working class, more than lower middle class. The most common one is middle class. Only 0,6% of people consider themselves higher class. The least supportive class is working class, and as the class level increases so does the support.

```{r, fig.height=6, fig.width=12}

# sd3 - Religion
prop.table(table(recoded_data$sd3, recoded_data$qc19), margin = 1) * 100

# sd1_7 - Friend/acquintance who is trans
prop.table(table(recoded_data$sd1_7))
prop.table(table(recoded_data$sd1_7, recoded_data$qc19), margin = 1) * 100

# qc2_recode  - Experienced discrimination
prop.table(table(recoded_data$qc2_recode, recoded_data$qc19), margin = 1) * 100

# school_group  - Age finished education
prop.table(table(recoded_data$school_group, recoded_data$qc19), margin = 1) * 100

# d63  - Subjective social class
prop.table(table(recoded_data$d63))
prop.table(table(recoded_data$d63, recoded_data$qc19), margin = 1) * 100

variables <- c("age_group", "polintr", "qa1", "d25", "d10")

# Lista para almacenar los gráficos
plots <- list()

# Generar gráficos para cada variable
for (variable in variables) {
  plot <- ggplot(recoded_data, aes_string(x = variable, fill = "qc19")) +
    geom_bar(position = "dodge") +
    labs(x = paste("Variable", variable), y = "Conteo", fill = "Variable cq19") +
    theme_minimal()
  plots[[variable]] <- plot
}

# Mostrar los gráficos en una cuadrícula
cowplot::plot_grid(plotlist = plots, nrow = 3, ncol = 2)
  
```

### Countries descriptive analysis

-   geo - By country. We observe that the least suporting country is Hungary with a 84% of people against, meanwhile the most supporting one is Spain.

-   Marriage. 42% of the European countries allow homosexual marriage. From countries which don't, there's a majority of people (59%) who are against trans people documentation change. However, in countries which do allow homosexual marriage, have a majority of 76% in favor.

-   Gini coefficient value. At first glance, there doesn't seem to be a clear tendency. Visually we can't detect right now wheter a higher or lower gini score is related to more or less support.

-   Hate crime laws. More than half of the European countries have hate crime laws (54%). In this case, both groups of countries have a similar response ratio, with a 3% more of people being in favor of our target question in the countries with these laws.

-   Banned conversion therapies. Only a 3% of countries had banned them in 2019. There's a big difference between both groups. In countries were they've been banned, a 85% of people agrees with allowing trans people change their documentation.

-   Existence of gender recogition administrative procedures. They are present in a 81% of countries. In countries where they don't exist, a 62% of people are against; meanwhile, a 60% of people are on favor on countries where they do.

-   No medical diagnosis needed for documentation changes. Only a 30% of countries allowed this in 2019. In countries where they allow it, a 79% of people is in favour, compared to the 48% of countries where they don't.

-   Dephatologisation. Only 6% of countries had specific laws regarding the dephatologisation of trans identities. In these few countries there's also a much higher acceptance: 80% in favour.

It's easy to tell that anti-discrimination laws are reflected in an increase in the acceptance of transgender individuals in the population of these countries.

```{r, fig.height=6, fig.width=12}

# By country
ggplot(data = recoded_data, aes(x = geo, fill = factor(qc19))) +
  geom_bar(position = "fill")+
  labs(x = "País", y = "Proporción", fill = "Valor en qc19") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  

prop.table(table(recoded_data$geo, recoded_data$qc19), margin = 1) * 100

# Marriage
prop.table(table(countries$Marriage))
prop.table(table(recoded_data$Marriage, recoded_data$qc19), margin = 1) * 100

# Gini value
data <- as.data.frame.table(prop.table(table(recoded_data$OBS_VALUE, recoded_data$qc19), margin = 1) * 100)
ggplot(data = data, aes(x = Var1, y = Freq, color = Var2, group = Var2)) +
  geom_line() +
  labs(x = "OBS_VALUE", y = "Proporción (%)", color = "Valor en qc19") +
  scale_color_manual(values = c("No" = "blue", "Yes" = "red")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Hate Crime Laws`
prop.table(table(countries$`Hate Crime Laws`))
prop.table(table(recoded_data$`Hate Crime Laws`, recoded_data$qc19), margin = 1) * 100

# Banned Conversion Therapies
prop.table(table(countries$`Banned Conversion Therapies`))
prop.table(table(recoded_data$`Banned Conversion Therapies`, recoded_data$qc19), margin = 1) * 100

# Existence of gender recogition administrative procedures
prop.table(table(countries$`Existence of gender recogition administrative procedures`))
prop.table(table(recoded_data$`Existence of gender recogition administrative procedures`, recoded_data$qc19), margin = 1) * 100

# No medical diagnosis needed
prop.table(table(countries$`No medical diag0sis needed`))
prop.table(table(recoded_data$`No medical diag0sis needed`, recoded_data$qc19), margin = 1) * 100

# Depathologisation
prop.table(table(countries$`Depathologisation`))
prop.table(table(recoded_data$`Depathologisation`, recoded_data$qc19), margin = 1) * 100


variables <- c("Marriage", "`Hate Crime Laws`", "`Banned Conversion Therapies`", "`Existence of gender recogition administrative procedures`", "`No medical diag0sis needed`", "`Depathologisation`")

# Lista para almacenar los gráficos
plots <- list()

# Generar gráficos para cada variable
for (variable in variables) {
  plot <- ggplot(recoded_data, aes_string(x = variable, fill = "qc19")) +
    geom_bar(position = "dodge") +
    labs(x = paste("Variable", variable), y = "Conteo", fill = "Variable cq19") +
    theme_minimal()
  plots[[variable]] <- plot
}

# Mostrar los gráficos en una cuadrícula
cowplot::plot_grid(plotlist = plots, nrow = 3, ncol = 3)


```

## Modeling

In this part, we will build multivariate logistic regression models to estimate the effects our predictors have on the outome variable. We have included mixed-effect models, in which the random effects will be countries, age groups, levels of educational attainment, religion and city type.

Our outcome variable, `qc19,` measures the odds/probability that an individual is in favor of changing the laws to allow transgender people the right to change their civil documents. In other words, this variable can be used to measure **transphobia**, which is defined as the dislike of or prejudice against transgender people.

Here are the various predictor variables we have chosen to include in the models. The data were taken from the 2019 survey conducted by the European Union.

-   life satisfaction-- factor `d70`

    -   "Very," "Somewhat," "Not really," or "Not at all satisfied"

-   politcial interest-- factor `polintr`

    -   "Frequently," "Occasionally," or "Never" talk about politics

-   opinion on globalization-- factor `qa1`

    -   Globalization is benefitting me "a lot," "somewhat," "not really," or "nor at all"

-   

### Split training/testing data

We split our data into a training and testing set. The training set will get 70% of data. This will allow us to test our model accuracy later with 'new' data.

```{r splitting, eval=FALSE}
library(caret)
# not sure if we need to split the data for surveys but anyway i will put it here just in case 
in.train <- createDataPartition(recoded_data$qc19, p = .7, list = F)
training.data <- recoded_data[in.train[,1], ]
testing.data <- recoded_data[-in.train[,1], ]
nrow(training.data)
nrow(testing.data)
```

### Three binomial logistic models

According to all 3 models below, **life satisfaction** has a strong effect on an individual's opinion about transgender rights. The model results show that respondents who said that they are "not very" and "not at all" satisfied with their life are much more likely to answer "No" to the question about giving transgender people the right to change their civil documents. The coefficients show that both negative responses have a similar effect on the odds that they will answer Yes/No. Respondents who answered "somewhat satisfied" are also more likely to answer "No" but not by much.

Other predictors **Age**, **Political Interest**, and **Country** also show a strong effect on the outcome.

testing

Including all our variables into the model, we observe many aren't significant. We reduce the variables until we select only those which are.

```{r}
modelo <- glm(qc19 ~ d70 + qa1 + polintr + d63 + d72_1 + d72_2 + sd3 + sd1_7 + d25 + d8 + qc1_4 + qc1_8 + qc6_1r + qc6_2r + qc6_3r + qc6_10r + qc7 + qc9_1 + qc9_4 + qc9_5 + qc9_10 + qc5_1 + qc5_3 + qc12_1r + qc12_2r + qc12_6r + qc12_10r + qc12_11r + qc15_1 + qc15_2 + qc15_3 + qc17_1 + qc17_2 + qc17_3 + qc17_3 + qc17_5 + qc18_1r + qc18_2r + qc18_3r + qc20 + d1 + qc2_recode + sd1_7 + d25 + Marriage + `Hate Crime Laws` + `Banned Conversion Therapies` +  `Existence of gender recogition administrative procedures`+ `No medical diag0sis needed` +  `Depathologisation`, data = recoded_data, family = binomial)


threshold = 0.6 # set the threshold here 
probabilities <- predict(modelo, newdata = recoded_data, type = "response")
predictions <- as.factor(ifelse(probabilities >= threshold, "Yes", "No"))
confusionMatrix(predictions, recoded_data$qc19)
```

In model we only keep the variables which are significative. From the variable qc9, the only

```{r}
modelo <- glm(qc19 ~ d70 + qa1 + d63 + sd3 + sd1_7 + d8 + qc1_4 + qc1_8 + qc6_1r + qc6_2r + qc6_3r + qc6_10r + qc7 + qc9_1 + qc5_1 + qc12_10r + qc15_1 + qc15_2 + qc15_3 + qc17_2  + qc17_5 + qc20 + d1 + d11 + d8 + qc2_recode + sd1_7 +  `Hate Crime Laws` +  `Existence of gender recogition administrative procedures`+ `No medical diag0sis needed` +  `Depathologisation` + OBS_VALUE, data = recoded_data, family = binomial)

summary(modelo)

threshold = 0.5 # set the threshold here 
probabilities <- predict(modelo, newdata = recoded_data, type = "response")
predictions <- as.factor(ifelse(probabilities >= threshold, "Yes", "No"))
confusionMatrix(predictions, recoded_data$qc19)

```

```{r models}

model1 <- glm(qc19 ~ d11 + d70*polintr,
             data = recoded_data, family = binomial)
model2 <- glm(qc19 ~ geo + sd3, 
             data = recoded_data, family = binomial)
model3 <- glm(qc19 ~ geo + d70 -1,
              data = recoded_data, family = binomial)
model4 <- glm(qc19 ~ OBS_VALUE + d70 + age_group + polintr + as.factor(qa1),
              data = recoded_data, family = binomial)

model5 <- glm(qc19 ~ OBS_VALUE*sd3,
              data = recoded_data, family = binomial)

model7 <- glm(qc19 ~ d70 + qa1 + d63 + sd3 + sd1_7 + d8 + qc1_4 + qc1_8 + qc6_1r + qc6_2r + qc6_3r + qc6_10r + qc7 + qc9_1 + qc5_1 + qc5_3 + qc12_10r + qc15_1 + qc15_2 + qc15_3 + qc17_2  + qc17_5 + qc20 + d1 + d11  + qc2_recode + + Marriage + `Hate Crime Laws` + `Existence of gender recogition administrative procedures` + `No medical diag0sis needed` + Depathologisation + OBS_VALUE, 
              data = training.data, family = binomial
  
)

stargazer::stargazer(model1, model2, model3, model4, type = "text")
exp(model7$coefficients)
```

```{r}
new.data <- data.frame(d11 = c(20, 30, 40),
                       d70 = c("Very", "Very", "Very"),
                       polintr = c("frequently", "frequently", "frequently"))
predict.glm(model1, newdata = new.data, type = "response")
```

### Testing accuracy of non-mixed effects models

With the threshold set at 50%, accuracy is only OK. Sensitivity is bad because there are lots of false positives. Most of the real "No"s are predicted as "Yes."

```{r}

threshold = 0.5 # set the threshold here 
probabilities <- predict(model7, newdata = testing.data, type = "response")
predictions <- as.factor(ifelse(probabilities >= threshold, "Yes", "No"))

confusionMatrix(predictions, testing.data$qc19)
```

```{r eval=true}
roc <- pROC::roc(testing.data$qc19, probabilities, na.rm = TRUE) 
sens <- roc$sensitivities #include sensitivities in test data
spec <- roc$specificities

plot(roc, col='green', print.thres=TRUE)
roc$levels

threshold = .577 # set the threshold here 
probabilities <- predict(mixed_model4, newdata = recoded_data, type = "response")
predictions <- as.factor(ifelse(probabilities > threshold, "Yes", "No"))

confusionMatrix(predictions, recoded_data$qc19)
```

### Mixed models

```{r}
library(lme4)

mixed_model1 <- glmer(qc19 ~ d11 + d70*polintr + (1|geo) ,
             data = recoded_data, family = binomial,
             )
mixed_model2 <- glmer(qc19 ~ d11 + d70 + polintr+ (1|geo), 
             data = recoded_data, family = binomial)
mixed_model3 <- glmer(qc19 ~ d70 + (1|geo),
              data = recoded_data, family = binomial)
mixed_model4 <- glmer(qc19 ~ d70 + (OBS_VALUE|sd3) +
                        polintr + qa1 + (d11|geo) + d10,
              data = recoded_data, family = binomial)

mixed_model5 <- glmer(qc19 ~ d10 + OBS_VALUE + (OBS_VALUE|sd3),
              data = recoded_data, family = binomial)
# reversing the order of levels for knowing a trans person. No is base.
recoded_data$know_transperson <- factor(recoded_data$sd1_7, levels = c("No", "Yes") )
levels(recoded_data$know_transperson)
mixed_model6 <- glmer(qc19 ~ d70 + qa1+ polintr + d72_1 + d72_2  + d10 +
                        know_transperson + d25 + d8 + d25 + (1|sd3) + (1|geo), 
              data = recoded_data, family = binomial
)

mixed_model7 <- glmer(qc19 ~ d70 + qa1 + d63 + (1|sd3) + sd1_7 + d8 + qc1_4 + qc1_8 + qc6_1r + qc6_2r + qc6_3r + qc6_10r + qc7 + qc9_1 + qc5_1 + qc5_3 + qc12_10r + qc15_1 + qc15_2 + qc15_3 + qc17_2  + qc17_5 + qc20 + d1 + d11  + qc2_recode + + Marriage + `Hate Crime Laws` + `Existence of gender recogition administrative procedures` + `No medical diag0sis needed` + Depathologisation + OBS_VALUE, data = recoded_data, family = binomial
  
)

stargazer::stargazer(modelo, mixed_model4, mixed_model5, type = "text", model.names = T)
summary(mixed_model7 )
stargazer::stargazer(confint.merMod(mixed_model7, method = "Wald"), type = "text")
```

```{r}

```

`mixed_model4` --

The 4th mixed model shows life satisfaction, political interest, sex, and opinion on globalization all have an effect on the outcome. The coefficient estimate for Intercept is not significant, probably because of the random effects in the model. Also, the life satisfaction factor variable for somewhat satisfied does not have a significant coefficient, indicating that given the other variables in the model, people who are somewhat satisfied with life have similar a response to those who are very satisfied. Very satisfied is the base value for this variable.

The results from the random effects show that the grouping variable *religion* has an effect on the intercept but only a minor effect on the slope for country-level income inequality. In other words, Atheists are much more likely to say "yes" for the outcome variable than other religious assocations. In countries that are more unequal, they are less likely than in countries that are more equal financially. But that slope is very small. For an increase of one Gini "score," the decline in odds of saying "yes" is only 0.9 percent. On the other hand, Sunni Muslims are much *less* likely to say "yes" than the average.

#### Visualizing the mixed models

```{r}
library(sjPlot)
sjPlot::plot_model(model2, type = "pred", 
                   terms =  "sd3", ci.lvl = 0.95) +
  theme_classic() +
  labs(
    x = "Religion",
    y = "Probability of Yes answer",
    subtitle = "Given their nationality, Sikhs have largest range at 95% confidence level "
  )


sjPlot::plot_model(mixed_model7, type = "pred", 
                   terms =  "sd1_7", ci.lvl = 0.95) +
  theme_classic() +
  labs(
    x = "Do I know a trans person?",
    y = "Probability of Yes answer",
    subtitle = NULL
  )

```

#### Testing accuracy of mixed models

```{r}
library(caret)
threshold = 0.6 # set the threshold here 
probabilities <- predict(mixed_model7, newdata = testing.data, type = "response")
predictions <- as.factor(ifelse(probabilities >= threshold, "Yes", "No"))

confusionMatrix(predictions, recoded_data$qc19)
```

```{r}
roc <- pROC::roc(recoded_data$qc19, probabilities, na.rm = TRUE) 
sens <- roc$sensitivities #include sensitivities in test data
spec <- roc$specificities

plot(roc, col='green', print.thres=TRUE)
roc$levels

threshold = .577 # set the threshold here 
probabilities <- predict(mixed_model4, newdata = recoded_data, type = "response")
predictions <- as.factor(ifelse(probabilities > threshold, "Yes", "No"))

confusionMatrix(predictions, recoded_data$qc19)
```

```{r}
cor.test((as.numeric(recoded_data$qc19)-1),recoded_data$OBS_VALUE)

recoded_data |>
  filter(!is.na(qc19)) |> 
  ggplot(aes(x = OBS_VALUE, y = sd3, size = qc19)) +
  geom_point()

recoded_data |> 
  group
table(recoded_data$sd3, recoded_data$qc19, )
```

## 

-   variables used in model - alicia

    -   why we picked them

    -   imputation process

-   descriptive analysis

    -   charts

-   explain the model and fixed effects of each parameter -giorgio

-   predictive model -eric

    -   random effect

    -   accuracy, ROC curve, ...

-   Conclusions! -eric

    -   what were we most surprised by

## Model 7 - Giorgio

```{r}
summary(model7)
roc <- pROC::roc(testing.data$qc19, probabilities, na.rm = TRUE) 
sens <- roc$sensitivities #include sensitivities in test data
spec <- roc$specificities

exp(model7$coefficients)
plot(roc, col='green', print.thres=TRUE)
roc$levels

confusionMatrix(predictions, testing.data$qc19)

```

Now about three statements regarding LGBT equality:


- qc5_1. Shared content online about discrimination incidents. People who have done this in the last 12 months seem to have a lower chance of supporting the target question. This might be related to supporting some other kinds of discrimination that aren't LGBT related.

- qc5_3 - Joined an association or campaign agains discrimination. People who have joined an association in the last 12 months are 12% more likely to answer positively to the target question.

- qc12_10r - How comfortable would you be with a gay, lesbian or bisexual coworker. When people answer they would be more comfortable, on a scale from 1 to 3, each 1 unit increase means a 8% increase in the chance of supporting trans people documentation change. This is the only variant of the question that was significant for the model.

- qc15_1 - Gay, lesbian and bisexual people should have the same rights as heterosexual people. When people answer higher values in the scale (disagree, totally disagree, 1 unit increase in the variable) the probability of answering positively to the target question lowers by -27%.

- qc15_2 - There's nothing wrong in same-sex relationships. As people answer more in disagree, the chance of supporting the target variable decreases.

- qc15_3 - Same sex marriage should be allowed in Europe. This question is a scale. Each 1 unit increase in the scale (from totally agree to agree, to disagree and totally disagree) is a 10% increase on people answering negatively to the target question. However this one is the statement with less impact in the model.

- qc17_2 - School should educate about ethnic origin or skin colour. The answers are in a scale (1: totally agree, 4: totally disagree). As people disagree more, there's a 14% decrease in the chance of them supporting trans people documentation change. 

- qc17_5 - School should educate about being intersex. When people disagree more with this statement, the probability of supporting the target question decreases. 

- qc20 - Believe that documents should have a third option for gender. When people answers no to that question, it's a 83% chance decrease in people being supportive of trans people changing their documents. 

- d1 - Ideology. As people identify more to the right, the support decreases. 1 unit increase in the ideology scale is a 3% decrease in support for trans people changing their documentation.

- d11- Age. As age increases, there's a slight increase in supportiveness. However in the descriptive analysis we saw how the oldest people were the less supportive. This might be due to people being more supportive in middle age but then decreasing in older ages.

- qc2_recode - Experienced discrimination. People who have experienced discrimination is 13% less likely to be supportive. This might be due to discrimination linked to certain social groups who are less likely to be supportive. Discrimination for experiencing LGBTphobia specifically wasn't significant for the model. 

- Legalised homosexual Marriage. Population from countries where there's legal homosexual marriage is slightly more likely to support trans people documentation change.

- Hate Crime Laws. In countries where there are hate crime laws, it's a bit less likely for the population to be supportive.

- Existence of gender recognition procedures. People who live in countries with these kind of laws are 27% less likely to support trans people. 

These results might be becase many countries have these laws and it doesn't seem to be linked to the support of the population. 

However, in countries with more progressive laws with have a focus and recognition of the problems trans people face, population seems to be more supportive:

- No medical diagnosis needed. People who live in countries where trans people don't need medical diagnosis to follow certain procedures are 1.8 times more likely to be supportive.

- Depathologisation laws. People who live in countries with specific dephatologisation laws towards transgender identities are 2.26 times more likely to answer positively to our target question than those who don't. However only 2 countries have these kind of laws in 2019.

And lastly, for the GINI index:

- OBS_VALUE - GINI index value. People who live in countries with a higher GINI index value are sllightly more likely to be supportive. Per each 1 unit increase in the index value, someone is 1% more likely to support trans people changing their documentation.